# Delayed trophic response of a marine predator to ocean condition and prey availability during the past century

##Abstract
Understanding the response of predators to ecological change at multiple temporal scales can elucidate critical predator-prey dynamics that would otherwise go unrecognized. We performed compound-specific nitrogen stable isotope analysis (CSIA) of amino acids on 153 harbor seal museum skull specimens to determine how this marine predator has responded to ecosystem change over the past century. The relationships between harbor seal trophic position, ocean condition, and prey abundance, were analyzed using hierarchical modelling of a multi-amino acid framework and applying 1-, 2-, and 3- year temporal lags. We identified delayed responses of harbor seal trophic position to both physical ocean conditions (upwelling, sea surface temperature, freshwater discharge) and prey availability (Pacific hake, Pacific herring and Chinook salmon). However, the magnitude and direction of the trophic response to ecological changes depended on the temporal delay. For example, harbor seal trophic position was negatively associated with summer upwelling, but had a 1- year delayed response to summer sea surface temperature, indicating some predator responses to climate extremes are not immediately observable. These results highlight the importance of considering dynamic responses of predators to their environment as multiple ecological factors are often changing simultaneously and predator response occurs at multiple temporal scales.

##Introduction
The regulation of food web structure by resources (bottom-up control) and the presence of top predators (top-down control) is fundamental for understanding food web responses to environmental, ecological, and anthropogenic change [@Carpenter1985; @Hunter1992; @Estes1998]. Ecological communities are continuously experiencing both biotic and abiotic disturbances [@Paine1998] and the ability of food webs to dynamically respond to these changes is crucial for ecosystem stability [@Ghedini2015]. In marine food webs, physical ocean conditions can impact primary production and ultimately constrain energy availability and thus biomass at higher trophic levels [@Ware2005; @Chassot2010; @Moore2018]. Similarly, the removal of top predators from an ecosystem as a result of human activities such as fishing can decrease predation pressure and alter abundance in both adjacent and non-adjacent trophic levels [@Heithaus2008; @Steneck2012]. However, large-scale changes in nutrient availability [@Rykaczewski2010], primary productivity [@Chassot2010], and top predator abundance over the past century [@Magera2013] means many food webs are experiencing shifts in multiple mechanisms of regulation in tandem, making it challenging to identify dominant drivers structuring ecosystems over the long term. 

Marine predators respond to multiple types of bottom-up drivers (i.e., ocean condition, prey availability) and the different temporal scales over which they respond is crucial for understanding community stability. However, delayed predator responses to environmental perturbations are prevalent in marine system, as impacts do not immediately propagate through the complete food web [@Duguid2019; @Smith2017]. Given communities can shift from bottom-up to top-down control, particularly in response to changing climate conditions [@Kratina2012], delayed predator responses to climate conditions has implications for abundance and mortality rates of prey.

Historical marine predator data that span multiple environmental, ecological, and anthropogenic contexts are useful for identifying time scales over which predators respond to ecosystem drivers. Compound-specific stable isotope analysis (CSIA) of amino acid nitrogen can serve as a tracer of historical predator response to ecological and environmental change by deriving retrospective trophic position estimates from museum specimens [@McMahon2019; @Feddern2021]. Source amino acids (i.e., phenylalanine, lysine, methionine) exhibit minimal trophic discrimination (the difference in ^15^N/^14^N between trophic and source amino acids in consumers from a trophic transfer) and thus are a proxy for the isotopic signature of primary producers at the base of the food web. In contrast, trophic amino acids (i.e., alanine, glutamic acid, valine, proline) demonstrate trophic enrichment [@McMahon2016] that varies for individual amino acids. Combined, this approach allows for reconstruction of historic trophic position estimates under changing environmental conditions when characterizing the isotopic baseline of past ecosystems may not be possible [@McMahon2019]. Thus, CSIA is well suited to identify long-term drivers of food web dynamics when analyzed with historic indices of ocean condition and prey availability.

Reconstructing time series of predator trophic position requires careful consideration of physiological and ecological parameters that contribute to stable isotope signatures. First, taxa exhibit different trophic enrichment factors based on excretion pathways, diet type (omnivory, herbivory, carnivory), and growth [@Nielsen2015]. Second, the nitrogen production pathway of vascular (i.e., seagrasses) versus nonvascular (i.e., marine diatoms) primary producers impart distinct stable isotope fractionation factors ($\beta$) as inorganic sources of nitrogen are converted to tissues [@VanderZanden2013]. Assumptions about the relative contributions of vascular versus nonvascular plants can therefore impact trophic position estimates (Choi et al. 2017). Finally, there is a delay between the time a prey source is consumed and when that prey source has been fully assimilated into the consumer, referred to as the ‘turnover time’. Turnover times must be considered when comparing trophic position data to ocean condition and prey availability covariates, as the consumer response to an ecological change will not be immediately observable in consumer tissues.

Nearshore coastal ecosystems provide a model system to assess long-term changes of food web drivers using archival museum specimens of a marine predator by applying CSIA. Food webs of coastal Washington and the Salish Sea have experienced dramatic restructuring over the past century due to declines and subsequent recoveries of marine predators [@Jeffries2003; @Ohlberger2019]. Decades of state-financed population control programs resulted in harbor seals (*Phoca vitulina*) reaching a historic low in the 1970's, with an estimated abundance of approximately 1,000 individuals [@Jeffries2003]. Following the cessation of bounties in 1960 and the passage of the Marine Mammal Protection Act in 1972, top-predator abundance increased dramatically. Benefitting from a relatively short life history, generalist diet, and legislation restricting mortality, harbor seal populations increased 10-fold between 1970 and 2003 [@Jeffries2003]. The dramatic increase in abundance of this top predator has been implicated in the declines in economically and ecologically important prey species in the region [@Chasco2017; @Nelson2019], specifically, Chinook salmon (*Oncorhynchus tshawytscha*). Chinook salmon are listed as endangered in the region (WDFW 2017) and are an important prey species for the endangered southern resident orca [@Marshall2015]. Simultaneously, the region has also experienced changes in nutrients (Mohamedali et al. 2011), climate regimes [@Corwith2002; @Mantua2002] and abundances of other important prey species such as Pacific herring (*Clupea pallasii*, @Siple2016). 

Here we examined a century of harbor seal trophic position data in coastal Washington and the Salish Sea. The objective of this work is to identify the time scales at which physical ocean conditions and prey availability exert bottom-up control on marine predator trophic ecology. We assumed a correlation between trophic position and prey species abundance is the result of increased or decreased consumption of that species. Additionally, we established a multi-amino acid framework for measuring trophic position that improves precision and ecological accuracy by applying a species-specific trophic discrimination factor [@McMahon2019; @Nielsen2015]. We also included a system specific $\beta$ value rather than a universal value, and applied temporal lags to account for both physiological and ecological delays in consumer response.

##Methods
###Sample collection and analysis
Samples were obtained using methods described in @Feddern2021. Briefly, harbor seal bone was obtained from four museum institutions (the Burke Museum, the Slater Museum, the Royal British Columbia Museum, and the Smithsonian Institute) and the National Marine Mammal Laboratory (NOAA). Specimens were treated by maceration in warm water and stored in acid free boxes. Sampling targeted adult specimens and prioritized long-term temporal coverage in two main regions: coastal Washington and the Salish Sea (which included 18 specimens from British Columbia). Specimens with sex, length, and age data were also prioritized but this information was not available for all sampled specimens. A total of 153 specimens were sampled with field collection dates ranging 1928-2014 (Figure \@ref(fig:map3)). 

###Trophic position determination
Bone collagen was decalcified, acid hydrolyzed, derivatized, and analyzed for nitrogen CSIA ($\delta^{15}N$) of 12 individual amino acids. Collagen samples were measured in triplicate with a laboratory standard containing a 12 amino acid mixture of known isotopic composition and a linear drift correction was applied. Full analytical details are described in Appendix 1: Text 1. Previous controlled feeding studies have determined the trophic enrichment factor (TEF) for harbor seals is substantially lower than the conventional literature value of 7.6‰ [@Germain2013] and thus applying a harbor seal-specific TEF is more accurate [@McMahon2015]. Therefore, trophic position was calculated using a harbor seal-specific TEF, described by @McMahon2015 as a "multi-TEF" approach, using the following equation:  

\begin{equation} 
Trophic Position =   
  \frac{\delta^{15}N_{(i-o)} - TEF_{(i-o),j} - \beta_{(i-o),N}}{\overline{TEF}_{(i-o)}}+2
  (\#eq:TP3)
\end{equation} 

where $\delta^{15}N_i$ is the measured stable isotope composition of a trophic amino acid i in a sample and $\delta^{15}N_o$ is the stable isotope composition of a source amino acid o in a sample.$\delta^{15}N_{(i-o)}$ represents the total trophic enrichment that has occurred throughout the food web measurable from predator tissues. $TEF_{(i-o),j}$ is the trophic enrichment factor between trophic amino acid *i* and source amino acid *o* of a specific consumer *j* (in this study, harbor seals) which occurs when consumer *j* assimilates prey. $\beta_{(i-o),N}$ is the difference in enrichment between a specific trophic amino acid *i* and source amino acid o for non-vascular primary producers *N* that occurs when primary producers assimilate inorganic nitrogen (@Nielsen2015; Table \@ref(tab:tpparam)). $\overline{TEF}_{(i-o)}$ represents the mean trophic enrichment that occurs at other trophic levels in the food web, and is calculated from the mean difference between trophic amino acid i and source amino acid o across all consumers described in @Nielsen2015.

The $\beta$ parameter differs substantially between vascular and nonvascular primary producers (@Ramirez2021; Table \@ref(tab:tpparam)). In food webs that assimilate organic matter from both vascular and nonvascular plants, including many nearshore food webs, β will be intermediate. In addition to testing a value that represents nonvascular primary producers exclusively $\beta_{(i-o),N}$, we also applied a two-source mixing model using carbon stable isotope data similar to @Choi2017. This generates a $\beta$ that is weighted $\beta_{(i-o),NV}$ based on the contributions of both vascular and nonvascular plants specific to the Washington nearshore ecosystem by first calculating the percent contribution of vascular plants to the food web:

\begin{equation} 
  \% V =
    \frac{\delta^{13}C_{H} - \delta^{13}C_{N}}
    {\delta^{13}C_{V} - \delta^{13}C_{N}}/100
    (\#eq:perV)
\end{equation} 

where $\delta^{13}C_{H}$ is the mean observed $\delta^{13}C$ value for Washington harbor seals. $\delta^{13}C_{V}$ is the carbon stable isotope end member for vascular plants, v (-9.5 ‰, derived from seagrasses *Zostera spp.*); and $\delta^{13}C_{N}$ is the carbon stable isotope end member for nonvascular plants, n (-19.5‰, derived from phytoplankton). Carbon end members were specific to the Washington nearshore ecosystems [@Howe2015]. Percent V is the percent contribution of vascular plants to the food web in which harbor seals forage. This assumes the trophic enrichment of $^{13}C$ is generally negligible (0–1‰, @Deniro1978).  $\beta_{(i-o),NV}$ was then derived by:

\begin{equation} 
\beta_{(i-o),NV} = (\beta_{(i-o),V}*\%V) +
(\beta_{(i-o),N}*(1-\%V))
  (\#eq:betanv)
\end{equation}

where $\beta_{(i-o),N}$ is the enrichment between an individual trophic amino acid i and source amino acid o for aquatic phytoplankton and $\beta_{(i-o),V}$  represents the trophic enrichment of seagrass which are vascular plants (Table \@ref(tab:tpparam)). 

###Quantifying bottom-up drivers of foraging 
To identify the most important explanatory variables of ocean condition and prey availability on predator trophic position, we fit two sets of candidate models using a multi-amino acid (glutamic acid, aspartic acid, alanine, proline, valine) hierarchical model. We selected 12 putative explanatory variables based on the length of the time series and divided them *a priori* into our two categories of interest, ocean condition and prey availability, representing our expected primary forcing mechanisms (Appendix S1: Tables 3.2 & 3.4). We fit candidate models to the trophic position and covariate data, and the candidate model set included a null and location-only model (Tables \@ref(tab:ocmod) & \@ref(tab:prmod)). Location (Salish Sea or coastal Washington) was included as a factor in all candidate models except the null model. Due to the correlation between the multivariate El Niño Southern Oscillation index and the Pacific Decadal Oscillation only one of these covariates were included in a single model. All timeseries were standardized around a mean of 0 and standard deviation of 1. To avoid collinearity, no more than four covariates (including location) were included in an individual model.

@Nielsen2015 determined that the use of multiple amino acids improves estimates of trophic position. Therefore, we used multiple trophic amino acids *i* (alanine, glutamic acid, valine and proline) and one source amino acid o (phenylalanine) to calculate trophic position. We selected amino acids based on: their prevalence in previous studies to derive parameters for equation 1; tissue turnover time relative to the source amino acid, phenylalanine; and their concentrations in bone collagen. The hierarchical linear model took the following structure:

\begin{equation} 
y_t = \alpha_k + \boldsymbol\beta X_{t-d} +\epsilon, 
  (\#eq:hier3)
\end{equation}

where *y* represents harbor seal trophic position from year *t* and *k* represents four different trophic amino acids (factors) used to calculate trophic position included as a random effects. *X* is a matrix of continuous bottom-up drivers in year *t*. $\boldsymbol\beta$ is a vector of predicted effects (coefficients) of bottom-up drivers included in the model (Tables \@ref(tab:ocmod) & \@ref(tab:prmod)) on harbor seal trophic position, and *a* is the predicted trophic position when all included bottom-up drivers are at an average value (represented by 0) in the coastal region of Washington. The variable *d* is the temporal lag between a change in bottom-up drivers and when that change is reflected in harbor seal bone collagen. This lag can be due to both physiological (tissue turnover) or ecological effects (rate of propagation through the food web). Time (year, Appendix S1: Section S4, Fig. S6), sex, size (Appendix S1: Section S3, Fig. S4 & S5), and seasonality (month, Appendix S1: Fig. S1), were also considered as predictors of trophic position but no significant associations were identified and thus these parameters were not included in the hierarchical modeling (Appendix S1: Section S3). The best performing models for both of these approaches were selected using Akaike’s Information Criterion (Akaike 1973) with a correction for small sample size (AICc). Inclusion of predictors in the model with the most support is indicative of ecological parameters that alter harbor seal foraging ecology or food web dynamics. Additionally, magnitude and sign of the coefficients for included predictors can be interpreted as the degree of trophic change induced by consuming different species, life stages of species, or groups of species, caused by a given predictor. 

\clearpage

##Tables

**Table** \@ref(tab:tpparam): Trophic amino acid specific parameter values for $\beta$ and trophic enrichment factors (*TEF*) to test parameterization of trophic position calculations using multiple TEFs and $\beta$ values. The source amino acid (*o*) for all parameters was phenylalanine. 

```{r tpparam,echo=FALSE}
supp1 <- read.csv("data/Ch3/TableS2.csv")

colnames(supp1) <- c("Trophic Amino Acid (i)",
                           "$\\beta_{(i-o),N}$",
                            "$\\beta_{(i-o),V}$",
                           "$\\beta_{(i-o),NV}$",
                           "$TEF_{(i-o),j}$",
                           "$\\overline{TEF}_{(i-o)}$")
supp1 <- kableExtra:::kable_styling(knitr::kable(supp1, "latex", escape = FALSE,
caption = "Trophic position parameter values",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8)
 
kableExtra::column_spec(supp1, column=2:4, width = "7em")

```

\clearpage


\begin{landscape}
Table 3.2: Datasets used to test ocean condition as a bottom-up driver of harbor seal trophic ecology. Total number of models tested = 35.

```{r envdat,echo=FALSE}
supp1 <- read.csv("data/Ch3/TableS3.csv")
 supp1<-  kableExtra:::kable_styling(knitr::kable(supp1, "latex",col.names = gsub("[.]", " ", names(supp1)), 
                  caption = "Environmental Datasets",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8)
kableExtra::column_spec(supp1, column=c(2,4), width = "27em")
```
\clearpage

Table 3.4:  Datasets used to test prey availability as a bottom-up driver of harbor seal trophic ecology. Total number of models tested = 26.

```{r preydat,echo=FALSE}
supp1 <- read.csv("data/Ch3/TableS4.csv")
 supp1<-  kableExtra:::kable_styling(knitr::kable(supp1, "latex",col.names = gsub("[.]", " ", names(supp1)), 
                  caption = "Prey Datasets",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8)
kableExtra::column_spec(supp1, column=c(2,4), width = "27em")
```
\end{landscape}
\clearpage

**Table** \@ref(tab:ocmod): Full candidate model set (n = 35) for ocean condition modelling. The same candidate models were used for the physiological delay, 1-year ecological delay, and 2-year ecological delay models 

```{r ocmod,echo=FALSE}
supp1 <- read.csv("data/Ch3/TableS5.csv")


supp1 <- kableExtra:::kable_styling(knitr::kable(supp1, "latex", escape = FALSE,
caption = "Ocean Condition Candidate Models",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8)
 
kableExtra::column_spec(supp1, column=2:4, width = "7em")

```
\clearpage

**Table** \@ref(tab:prmod): Full candidate model set (n = 26) for prey availability modelling. The same candidate models were used for the physiological delay, 1-year ecological delay, and 2-year ecological delay models. 

```{r prmod,echo=FALSE}
supp1 <- read.csv("data/Ch3/TableS6.csv")


supp1 <- kableExtra:::kable_styling(knitr::kable(supp1, "latex", escape = FALSE,
caption = "Prey Availability Candidate Models",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8)
 
kableExtra::column_spec(supp1, column=2:4, width = "7em")

```


##Figures

**Figure** \@ref(fig:map3): Spatial distribution of harbor seal specimens (a) collected in the Salish Sea (yellow) and coastal Washington (blue) with the year of specimen collection and total number of specimens (n) for each year from 1928-2014 in the Salish Sea (b) and coastal Washington (c).
\newline 

\begin{figure}[h]
\centering
  \includegraphics[width=0.85\textwidth]{figure/Ch3/Figure1.pdf}
  \caption{Spatial and temporal distribution of specimens}
  (\#fig:map3)
\end{figure}

\clearpage
