<!--
The {#rmd-basics} text after the chapter declaration will allow us to link throughout the document back to the beginning of Chapter 1.  These labels will automatically be generated (if not specified) by changing the spaces to hyphens and capital letters to lowercase.  Look for the reference to this label at the beginning of Chapter 2.
-->

# The influence of temporal isotope heterogeneity and isotope incorporation rates on consumer trophic position estimation

##Abstract
Nitrogen stable isotope ratios are frequently used to estimate 	Nitrogen stable isotope ratios are frequently used to estimate trophic position of consumers. A key assumption in this application of stable isotope data is that the food webs in which consumers forage are in isotopic equilibrium with nitrogen resources. However, isotopic heterogeneity in the food web, which can arise from seasonal nutrient fluctuations, spatial variability in primary production, or variability in available prey, exists in most systems. Given resources are not immediately assimilated into the tissues of consumers, it is likely that the assumption of isotopic equilibrium is frequently violated in most systems. We tested the degree to which the violation of this assumption impacts consumer stable isotope ratios and trophic position estimates from those ratios. We constructed a compartment model to explore consumer stable isotope incorporation rates and their effect on trophic position calculations using bulk and compound-specific stable isotope data from previous experimental and observational studies derived from a variety of aquatic systems. We also tested how different parameterizations of the stable isotope baseline in trophic position calculations can improve accuracy. We found trophic position estimates of higher trophic level consumers are less accurate than lower trophic level consumers when applying bulk stable isotope analysis (BSIA) with particulate organic matter as the stable isotope baseline in trophic position equations. Accuracy of consumer trophic position improved for tertiary consumers when applying a 90-day lag between the consumer stable isotope measurement compared to baseline measurements, but this was at the expense of decreased accuracy for lower trophic level consumers. Compound-specific stable isotope analysis (CSIA) of individual amino acids was more accurate in estimating trophic position for all consumers and systems compared to BSIA. Overall, our results show consideration of stable isotope heterogeneity and stable isotope incorporation rates of consumers is important for accurate trophic position estimates and should be carefully considered when designing stable isotope studies.
\clearpage

##Introduction
In most ecosystems, seasonality drives food web structure and function [@McMeans2015]. Nutrients, such as phosphorus, nitrogen, and silicon, can be seasonally limited in their availability, which can determine the timing and amount of primary production [@Moon2007; @Kolzau2014; @Sondergaard2017] in aquatic systems. Nutrient availability can be influenced by oceanographic conditions, due to upwelling [@Barth2007; @Ferreira2020], resource pulses [@Benbow2020; @Verspoor2011], or anthropogenic inputs [@Khangaonkar2019; @McCrackin2017], allowing for increased resource sequestration as primary producers are released from growth-limiting constraints. In addition, the impact of nutrient limitation can propagate to the rest of the food web, as primary production can alter abundance in both adjacent and non-adjacent trophic levels [@Ware2005] by controlling the energy available for higher trophic level consumers.

Stable isotope analysis of nitrogen ($^{15}N/^{14}N$) serves as a useful tool for identifying trophic interactions and nutrient sources in food webs. The nitrogen stable isotope values ($\delta^{15}N$) of consumers within a food web are governed by nitrogen availability, contributions of distinct nitrogen sources (i.e., anthropogenically fixed; @VanderZanden2005), and trophic position. Trophic position of a consumer may be calculated based on marked isotope enrichment between consumer and prey (typically 2 to 4 per mille, @Deniro1981) referred to as the trophic enrichment factor. However, these calculations also require, and are highly sensitive to, the stable isotope baseline (stable isotope ratio of primary producers or inorganic nitrogen sources) within the food web, which can vary through space and time [@Post2002; @Anderson2007; @Possamai2021]. There is no consistent methodology to measure isotope baselines [@Kjeldgaard2021; @Possamai2021] although recommendations for best practices have been made [@Post2002; @Jardine2014; @Possamai2021]. Typically, the isotope baseline is estimated from primary producers, particulate organic matter (POM), or primary consumers that were measured at the same time as the consumer of interest [@Kjeldgaard2021]. However, baseline values averaged over time or generic values that may not be from the same environment or time period are also used (i.e. @Estrada2003; @RuizCooley2006; @Chiang2020] and fail to represent heterogeneity within and across systems.

A fundamental assumption of trophic position calculations from stable isotope data is that a consumer is in close equilibrium or ‘isotopic steady state’ with its dietary resources [@Deniro1981; @DelRio2009; @Phillips2014] and the isotope baseline used to calculate consumer trophic position. This means the stable isotope inputs (consumed resources) are in balance with outputs (excretion) such that consumer tissues represent the stable isotope value consumed resources plus a trophic enrichment factor. However, in nature such steady state do not necessarily occur, and while such assumption violations are often acknowledged, they are rarely quantified. The ecosystem heterogeneity created by seasonal fluctuations in both physical, chemical, and biological processes influence the isotopic values of primary producers and primary consumers (the isotope baseline) which varies both seasonally [@Woodland2012] and spatially [@Vokshoori2014]. For example, seasonal variability in the nitrogen isotope values of seston or particulate organic matter are common in aquatic systems [@Wu1999; @Rolff2000; @Syvaranta2006; @Matthews2007; @Gu2009; @Vokshoori2014]. Such isotope variation at the base of the food web has been detected in higher trophic level consumers [@Popp2007; @Olson2010; @Dale2011; @Feddern2021]. However, isotope incorporation into consumer tissues typically occurs at a slower rate than that of the prey, as the rate of isotope turnover increases with animal size [@Thomas2015; @VanderZanden2015] in addition to temperature, growth rate, and the type of tissue studied [@Dalerum2005; Figure \@ref(fig:TissueTurn)]. As a result of slower incorporation times, consumers can dampen short-term temporal stable isotope heterogeneity relative to algae or particulate organic matter [@Woodland2012]. The temporal mismatches created by differences in isotope incorporations rates between diet and consumer result in non-equilibria that violate stable isotope assumptions and can affect the calculation of animal trophic position. 

A recent methodological advancement in stable isotope analysis, compound-specific stable isotope analysis (CSIA) of amino acid nitrogen ($\delta^{15}N$), provides researchers with a powerful new tool to estimate the stable isotope values of the base of the food web [@Sherwood2011; @McMahon2015; @Feddern2021] which can be applied for trophic position calculations of a consumer [@Popp2007]. A key advantage of estimating trophic position using CSIA is the distinct properties of source and trophic amino acids. Source amino acids do not exhibit isotopic enrichment with each trophic transfer and thus depict the nitrogen stable isotope baseline of a system. In contrast, trophic amino acids track the apparent trophic position of the animal through trophic enrichment. This allows estimation of baseline values directly in the consumer organism of interest. Furthermore, the baseline value of the source amino acids represents an integrated dietary contribution relating directly to the physiology and thus stable isotope turnover of the consumer itself. That, in principle, results in a smaller temporal mismatch and standardization across a heterogenous environment between the source and trophic amino isotope values.

Incorporation rates of individual amino acids vary [@Bradley2014; @Downs2014; Figure \@ref(fig:TissueTurn)) and whether source and trophic amino acids are incorporated at similar rates is currently not well studied. @Downs2014 showed that the amino acid incorporation rates of glutamic acid (trophic) and phenylalanine (source), which are the most commonly applied amino acids for trophic position calculations due to their trophic enrichment factors, in shrimp appear similar (Figure \@ref(fig:TissueTurn)). However, it is not known if this is a persistent pattern between glutamic acid and phenylalanine across animal groups or tissue types because controlled feeding studies measuring turnover rates of individual amino acids are currently limited. There is evidence that turnover time of amino acids may also vary with diet quality for some but not all amino acids [@Chikaraishi2015]. Furthermore, the same study found both faster and slower turnover rates for other trophic (aspartic acid, proline) and source (lysine, methionine) amino acids, motivating an analysis of how temporal isotope integration in amino acids influences the accuracy of trophic position estimates. 

Here we evaluate the consequences of biologically-plausible consumer isotope incorporation rates under baseline variability for estimating trophic position using bulk stable isotope analysis (BSIA) and CSIA. We approach this by constructing a simulation model using stable isotope data from previous experimental and observational studies. The objectives of this work are to:

1. Explore how mismatches in isotope turnover rates between food web baseline and consumers influence trophic position calculations from bulk $\delta^{15}N$ values among various trophic levels, tissue types, and aquatic systems.

2. Examine how variability of source amino acid and trophic amino acid turnover rates can influence trophic position calculations.

3. Consider how different trophic position calculations and isotope baseline sampling strategies for both BSIA and CSIA can improve accuracy in trophic position estimations.

##Methods

Nitrogen stable isotope values of primary, secondary, and tertiary consumers were modelled using a first order-kinetics compartment model [@Cerling2007; @DelRio2012]. Different stable isotope baselines and incorporation rates were modelled to understand how they impact stable isotope values and trophic position accuracy of consumers for distinct food web scenarios. To represent biologically plausible parameters, stable isotope baselines were compiled from observational studies and tissue turnover times for BSIA and CSIA were compiled from controlled feeding experiments. Trophic position estimates were calculated from model simulations and accuracy of trophic position estimates were quantified for each consumer and each model simulation. 

###Tissue Turnover Modelling Structure 

The stable isotope values of consumers were modeled using a first order kinetics model as described by @DelRio2012 and modified from @Cerling2007. Many controlled feeding studies use an exponential fit model to estimate tissue turnover where:

\begin{equation} 
\delta X_t = be^{-\lambda t}+c
(\#eq:one)
\end{equation}

$\delta$ is the stable isotope value of a consumer’s tissues for a given element *X* that has been measured at time *t* before and after the consumer has been switched to an isotopically distinct diet. The parameters *b*, $\delta$, and *c* are each estimated empirically from a best fit of the exponential model to the data, where the parameter *c* is the asymptotic value of the isotope composition of tissue after the diet switch once tissues have reached steady state. The parameter *b* represents the magnitude of isotopic change in tissues or the difference between the initial isotopic steady state before the diet switch and the final isotopic steady state after the diet switch. The parameter $\lambda$ is a first order data-derived rate constant referred to as the turnover rate, which can be derive the isotopic half-life or time to reach steady state after a diet switch:

\begin{equation} 
  t_\alpha = \frac{ln(1-\alpha)}
  {\lambda}.
  (\#eq:two)
\end{equation}

In most controlled feeding studies $t_\alpha$ is reported as the isotopic half-life where $\alpha$ is 0.5 which represents the amount of time (in days or hours) required for 50\% equilibration with a new diet source. $t_{0.95}$ values are also frequently reported and denote 95\% equilibrium and are commonly accepted as the amount of time to reach isotopic steady state. Chemically, this representation of the first order rate constant, $\lambda$, assumes a first order reaction given for a trace isotope where the concentration of the rare isotope ($^{15}N$) is significantly less than that of the abundant isotope ($^{14}N$). Thus, the system can be closely approximated by changes in only the rare isotope. In the case of controlled feeding experiments the food source is treated as a reservoir where the isotopic value of the food supply is unaffected by consumption [@Criss1999].

We apply this approach but with a modified representation of equation \@ref(eq:one) as presented in @DelRio2012 with more biologically meaningful set of parameters: 

\begin{equation} 
  \delta X_t = \delta X_\infty - 
  (\delta X_\infty - \delta X_0)e^{\lambda t},
  (\#eq:three)
\end{equation}

where *x* represents a particular element (in this case nitrogen), $\delta X_\infty$ is the asymptotic isotope value a consumer will reach at steady state with a new diet (or the final diet isotopic value plus the trophic enrichment factor). $\delta X_0$ is the initial isotope value of the tissue, and $\lambda$ is the empirically derived rate constant as described above. 

###Bulk Stable Isotope Model

Using observed stable isotope baseline data derived from particulate organic matter (POM) (Table \@ref(tab:base)) we modelled the stable isotope composition of primary, secondary and tertiary consumers associated with a particular food web scenario (Figure 5.1). The number of trophic transfers is a known component of the model structure and for all model simulations the true trophic level was 2 for the primary consumer, 3 for the secondary consumer, and 4 for the tertiary consumer. Equation \@ref(eq:three) was modified to describe each trophic level as: 

\begin{equation} 
  \delta X_{t,\hat{tp}} = 
  (\delta X_{t-1,\widehat{tp}-1} + TEF) - 
  [(\delta X_{t-1,\widehat{tp}-1} + TEF) -
  \delta X_{t-1,\widehat{tp}}]e^{- \lambda},
  (\#eq:four)
\end{equation}

where, $\delta X_{t,\widehat{tp}}$ is $\delta^{15}N$ at time *t* with a true trophic position of $\widehat{tp}$ of a modelled consumer. *TEF* is the trophic enrichment factor of consumers which was assumed to be known and the canonical value of 3.4 per mille was applied for all consumers in the BSIA models. Trophic position was then calculated from modelled stable isotope date $\delta X_{t,\widehat{tp}}$ from equation \@ref(eq:four) as:

\begin{equation} 
  tp = \frac{\delta X_{t, \widehat{tp} = 2:4}-\delta X_{t,\widehat{tp} = 1}}
  {TEF} + 1,
  (\#eq:five)
\end{equation}

where *tp* is the estimated trophic position. $\delta X_{t,\widehat{tp}=1}$ represents the $\delta^{15}N$ of primary producers at the base of the food web and $\delta X_{t,\widehat{tp}=2:4}$ represents the modelled consumers (primary, secondary, and tertiary). Finally, an effect size was estimated as the magnitude and duration the estimated trophic position, *tp*, deviated from the true trophic position $\widehat{tp}$ (Figure \@ref(fig:EffectSize)):

\begin{equation} 
  d = tp - \widehat{tp},
  (\#eq:six)
\end{equation}

where *d* is the trophic position estimation error of a consumer based on the estimated trophic position *tp* calculated from equation \@ref(eq:five), where $\widehat{tp}$ represents the true trophic position ( $\widehat{tp}$ = 2:4) of primary, secondary, and tertiary consumers. Duration of deviation was determined by summing the number of days where *d* was ± 0.5 for each consumer (Figure \@ref(fig:EffectSize)), meaning *tp* was within 0.5 of the true trophic position, $\widehat{tp}$. We assumed the stable isotope signature of the base of the food web, *TEF*, and $\lambda$ were known. 

###Compound-specific Stable Isotope Model

Diet switching studies that report the isotope turnover rates for individual amino acids are currently uncommon in the stable isotope literature. For the CSIA model simulations we applied our modelling framework to two consumers with known isotope turnover rates for individual amino acids, bluefin tuna (*Thunnus orientalis*; @Bradley2014) and Pacific white shrimp (*Litopenaeus vannamei*; @Downs2014). For each consumer, four trophic amino acids (glutamic acid, alanine, proline, valine) and a single source amino were modelled (Figure \@ref(fig:TissueTurn)). The canonical source amino acid for trophic position calculations, phenylalanine, was modelled for shrimp however a phenylalanine half-life value was not reported in @Bradley2014 and lysine was used as the source amino acid for tuna instead of phenylalanine. Tuna and shrimp were modelled as a single food web (Figure \@ref(fig:CSIA)) for each of the four baseline case studies (Table 5.1) where shrimp was the primary consumer and tuna was the secondary consumer. Equation 4 was modified to model each of the 5 amino acids for each consumer:

\begin{equation} 
  \delta X_{t,\widehat{tp},a} = 
  (\delta X_{t-1, \widehat{tp}-1,a} + TEF) -
  [(\delta X_{t-1, \widehat{tp}-1,a}+TEF) - 
  \delta X_{t-1, \widehat{tp},a}]e^{-\lambda_a},
  (\#eq:seven)
\end{equation}

where $\delta X_{t,\widehat{tp},a}$ is the estimated stable isotope composition of nitrogen, *X*, at a given time *t* for a specific consumer,$\widehat{tp}$, for each amino acid *a*. As such, $\lambda$ was distinct for each amino acid, *a* including the source amino acid. The trophic position equation was also modified to calculate the trophic position of shrimp and tuna from amino acids, with source amino acids instead of the measured baseline:

\begin{equation} 
  tp_t = \frac{\delta X_{t, \widehat{tp} = 2:3, (i-o)}
  -\beta_{(i-o)}}
  {\overline{TEF}_{(i-o)}} + 1,
  (\#eq:eight)
\end{equation}

where trophic position tp for time *t* is estimated from simulated stable isotope data from equation \@ref(eq:seven). $\delta X_{t,\widehat{tp}=2:3,(i-o)}$  is the stable isotope value of compound *X* (nitrogen for this study) for each trophic amino acid *i* and source amino acid *o* in a consumer with a known trophic position $\widehat{tp}$̂. $\delta X_{i-o}$ represents the total trophic enrichment that has occurred throughout the food web measurable from consumer tissues at a given trophic level ($\widehat{tp}$). $\beta_{i-o}$ is the difference of enrichment between a specific trophic amino acid *i* and source amino acid *o* for primary producers derived from laboratory experiments (@Chikaraishi2007). $TEF_{i-o}$ represents the mean trophic enrichment factors for consumers, and is calculated from the mean difference between trophic amino acid *i* and source amino acid *o* across all consumers described in @Nielsen2015. Equation \@ref(eq:six) was applied to the CSIA models calculate trophic position deviation *d* with $\widehat{tp}$ = 2 for shrimp and $\widehat{tp}$ = 3 for tuna.

###Data Compilation

Data used for case studies was compiled from previous research that measured the stable isotope signature of nitrogen in particulate organic matter (POM) in both marine and freshwater systems and laboratory experiments that measured nitrogen stable isotope turnover rates in different taxa, tissues, and amino acids. These studies were used to obtain realistic estimates of the natural variation of δ15N at the base of the food web over the course of a year. We prioritized studies that 1) collected data at least monthly and 2) represented a range of aquatic systems (Table \@ref(tab:base)). Four diverse ecosystem studies were used as case studies; a moderately eutrophic urban lake [@Syvaranta2008], a shallow subarctic lake in interior Alaska [@Gu1994], a pristine oligotrophic lake [@Mathews2007] and a marine upwelling site near Vancouver Island [@Wu1999]. The rate of tissue turnover for nitrogen isotopes is typically measured on the time scale of hours or days and thus daily values of the nitrogen isotope baseline is necessary to model isotopic incorporation by consumers. However, studies collect stable isotope data from POM and primary producers on weekly and monthly scales. Therefore, we linearly interpolated nitrogen stable isotope values between observations in order to approximate daily changes in the isotopic baseline (Table \@ref(tab:base)).

Six theoretical food web scenarios were constructed for BSIA that contain primary, secondary, and tertiary consumers and that could be associated with the baseline values for each of the four case studies (Table \@ref(tab:base)& Figure 5.1). Data for bulk nitrogen stable isotope turnover rates are available for a variety of taxa and tissues [@VanderZanden2015].  Food web scenarios were constructed by matching consumers’ half-life values with isotope baseline data from an ecosystem in which they could reasonably forage. A marine food web was constructed from half-life values for an amphipod, a herring and a great skua and a freshwater food web was constructed from half-life values for a water strider, a steelhead or small mouth bass, and a black bear (Figure 5.1). The selected species did not represent a specific food web, but rather were selected based on whether they could reasonably be found within the same food web and if their tissue turnover times could be representative of other species within their trophic level. These scenarios were used to make specific comparisons regarding stable isotope baseline and the half-life values of the consumers. As a first comparison, the half-life value of the secondary consumer was increased between scenario a and b to represent different tissues (liver versus muscle), an increase in temperatures, or an increase in size, all of which can alter isotope incorporation rates. Comparison 2 decreased the half-life of the secondary consumer and increased the half-life of the tertiary consumer between scenarios c and d (Figure 5.1). The objective of this comparison was to identify how dietary differences of the tertiary consumer can impact its stable isotope composition and how sampling different tissues of the tertiary consumer can impact trophic position estimates. Finally, comparison 3 examined how two isotope basslines with distinct shapes can impact trophic position estimates of the same food web (Figure 5.1, Table \@ref(tab:base)). 

For CSIA of amino acid nitrogen models, half-life data from two aquatic consumers (shrimp and tuna) were used to construct a single food web with four trophic amino acids and one source amino acid modelled for each consumer (Figure \@ref(fig:CSIA)). The stable isotope values for each amino acid for both consumers were modelled for all four of the stable isotope baselines derived from particulate organic matter (Table \@ref(tab:base)). The stable isotope values for each amino acid were then compared for each baseline and both consumers to assess the performance of trophic amino acids in trophic position estimations. 

###Comparison to recommended methodologies

Guidance on baseline sampling for BSIA suggests primary consumers should be used as the stable isotope baseline rather than particulate organic matter or primary producers [@Post2002], and consumers of interest should be sampled 90 days after sampling the stable isotope baseline [@Possamai2021]. These two approaches can minimize baseline isotope variability or account for disconnect between the baseline and consumer due to tissue turnover. We assessed the utility of these approaches to improve trophic position accuracy across a variety of systems by applying them to our simulated food webs. To calculate trophic position using the primary consumer we applied:

\begin{equation} 
  tp = \frac{\delta X_{\widehat{tp}}
  -\delta X_{\widehat{tp}=2}}{TEF}+2.
  (\#eq:nine)
\end{equation}

Using the stable isotope values of the simulated primary consumer, $\delta X_{\widehat{tp}=2}$. To calculate the trophic position using a 90 day delay we used:

\begin{equation} 
  tp_y = \frac{\delta X_{\widehat{tp},y}
  -\delta X_{\widehat{tp}=1,y}}{TEF}+1,
  (\#eq:nine)
\end{equation}

where trophic position of consumers, *tp*, is estimated at days (*y*) = 90, 270 and 360 from the simulated consumer stable isotope data $\delta X_{\widehat{tp},y}$  for each consumer, $\widehat{tp}$. $\delta X_{\widehat{tp}=1,y}$ is the stable isotope value of the baseline at days, (*y*) = 1, 90, 270, and 360. A third approach combining both of these methodologies was also applied to the model output. Trophic position estimation error for all three approaches was assessed using *d* from equation \@ref(eq:six).

###Sensitivity Analysis
A sensitivity analysis was performed to further explore how variability in tissue turnover time impacts trophic position estimates using BSIA and CSIA. We simulated an isotope baseline that had a rate of isotope change that varied from 0 to 0.15 (per mille per day), represented 123 days, and represented a 1 to 6 per mille isotope change which was realistic based on previous studies (Table \@ref(tab:base), Figure \@ref(fig:BSIA15N)). A primary, secondary, and tertiary consumer were modelled for BSIA (equation \@ref(eq:four) - \@ref(eq:six)) and CSIA (equations \@ref(eq:seven) - \@ref(eq:eight)). Half-life values ranged from 1 to 280 days which were considered reasonable estimates for most tissues, taxa, and both approaches (Figure \@ref(fig:TissueTurn)). The model was run using the same simulated baseline for each of the 280 half-life values. To assess the sensitivity across half-life values, the mean and standard deviation of trophic position deviation was taken for each model run  which represented the average deviation across the 123-day simulation. For the CSIA model, two analyses were run, one with a source amino acid half-life value of 130 days (representing the values for lysine in bluefin tuna reported by @Bradley2014) and a second of 33 days (representing the values for phenylalanine in shrimp reported by @Downs2014). The half-life value of the source amino acid was assumed to be known and the half-life value of the trophic amino acid was varied for each simulation. The half-life values of the primary, secondary, and tertiary consumers were manipulated simultaneously, and all three consumers had the same half-life value for a single model simulation. 

##Results

The dynamic behavior of the stable isotope baseline impacts the BSIA values and estimation error of trophic position for simulated primary, secondary, and tertiary consumers. The four stable isotope baselines (Table \@ref(tab:base)) showed a range of stable isotope variability over the course of a year (Figure \@ref(fig:BSIA15N)). The oligotrophic lake (baseline 4, Figure \@ref(fig:BSIA15N)F) had the greatest variability in the stable isotope values of particulate organic matter, that ranged from 0 to over 10‰, but was relatively stable (fluctuations within 2‰ range) until day 300. In contrast, the two other freshwater systems exhibited fluctuations within a 5‰ range, but oscillated more frequently between values (Figure \@ref(fig:BSIA15N)C – E) within that range. Of the two eutrophic lakes, the heavily polluted lake (baseline 3; Figure \@ref(fig:BSIA15N)E) had the greatest range in stable isotope values. The marine upwelling system showed baseline isotope values characteristic of seasonal nutrient fluctuations and spring phytoplankton blooms. Nitrogen stable isotope values of POM were lowest in the spring and summer when upwelling is high, and higher in the winter when nutrients are more limiting (Figure \@ref(fig:BSIA15N)A & B).

###Bulk Stable Isotope Model

Our model showed that consumers are rarely in isotopic steady state with their available resources and the violation of this assumption resulted in trophic position estimation error by as much as one trophic level (Table \@ref(tab:durdev); Figure \@ref(fig:BSIATP)C & F). We define duration of deviation as the number of simulated days that a consumer had a trophic position estimation error of greater than ± 0.5 trophic level (equivalent to ± 1.7‰ deviation from steady state with resources). Primary consumers had less than 34 days duration of deviation (Table \@ref(tab:durdev)). The duration of deviation increased as you moved up the food web. For all scenarios, tertiary consumers had the longest duration of deviation, which was as high as 115 out of 365 days for scenario 4 (Table \@ref(tab:durdev); Figure \@ref(fig:BSIA15N)C). Based on comparison 1 (increasing half-life of the secondary consumer only), the duration of deviation also increased for tertiary consumers when half-life of secondary consumers increased, despite the same half-life of tertiary consumers and same stable isotope baseline (Table \@ref(tab:durdev); Figure \@ref(fig:BSIA15N)A versus B). Therefore, duration of deviation was determined by the consumers half-life, in addition to the half-life of consumers lower in the food web that serve as resources for higher trophic levels.

Similarly, the magnitude of trophic position estimation error depended on both the half-life of the consumer of interest, and the half-life of consumers lower in the food web. When a primary or secondary consumer is not in steady state with its resources, the resulting trophic position deviation propagates through the food web. In comparison 1, the magnitude of trophic position estimation error for the tertiary consumer was higher when the half-life value of the secondary consumer (Figure \@ref(fig:BSIATP)A) was higher, despite the half-life value of the tertiary consumer being the same in both scenarios. Sampling a tissue of a tertiary consumer with extremely rapid turnover time (i.e., blood plasma Figure \@ref(fig:BSIATP)C) did not prevent trophic position estimation error, despite the tertiary consumer being as close to isotopic steady state with its prey as possible. Sampling the blood plasma of a tertiary consumer that has prey with slow tissue turnover (i.e., adult steelhead muscle), results in substantial trophic position estimation error ranging from -1.1 to 0.8 ($\widehat{tp}$ = 4, Figure \@ref(fig:BSIATP)C). This deviation is similar to sampling a tissue with a longer isotopic half-life from a tertiary consumer that eats prey with a faster half-life (i.e., juvenile small mouth bass muscle) ($\widehat{tp}$ = 4, Figure \@ref(fig:BSIATP)5D) which ranges from -0.9 to 0.8. 

In addition to half-life value of the consumers in a food web, the dynamic behavior of the stable isotope baseline (rate of isotopic change, inflection points, magnitude of change) also impacts the stable isotope values of the consumers, and as a result, the magnitude and duration of trophic position estimation error. The eutrophic urban lake [@Syvaranta2008] had a stable isotope baseline that demonstrated a slow rate of change that reached a high magnitude (Figure \@ref(fig:BSIA15N)E). By comparison, the pristine oligotrophic lake showed frequent, low magnitude, rapid, fluctuations followed by a rapid increase of 10‰ (Figure \@ref(fig:BSIA15N)D). The eutrophic lake had a much longer duration of deviation for all consumers (Table \@ref(tab:durdev)) compared to the oligotrophic lake (Figure \@ref(fig:BSIATP)E & D). For almost the entire model simulation, the oligotrophic food web was within ± 0.5 trophic levels until day 310 when all consumers experienced a rapid, high-magnitude, deviation from their true values (Figure \@ref(fig:BSIATP)D). In comparison, the trophic position of the consumers in the eutrophic lake were consistently underestimated but experienced a lower magnitude of deviation (Figure \@ref(fig:BSIATP)E).

###CSIA Model

The magnitude of trophic position estimation error and the duration of deviation was lower using CSIA compared to BSIA and varied across trophic amino acids. Baseline 3, the moderately eutrophic urban lake, had the highest magnitude of estimation error for all stable isotope baselines for the CSIA model (Figure \@ref(fig:CSIATP)E & F) but did not exceed ± 0.4 estimation error for primary or secondary consumers. In contrast, for the BSIA model secondary consumers exceeded ± 0.5 estimation error for all scenarios (Figure \@ref(fig:BSIATP)). For the BSIA model, estimation error always propagated up the food web and was greatest for higher level consumers in all scenarios. This was not necessarily the case for the CSIA model. In fact, for most baselines and amino acids, both shrimp and tuna performed similarly (Figure 7) despite tuna being modelled as a higher trophic level consumer (secondary) than shrimp (primary). No amino acids for the primary and secondary consumers in the CSIA model exceeded an estimation error of greater than 0.4 trophic levels and for most trophic amino acids it did not exceed ± 0.2 trophic levels, with glutamic acid and alanine preforming particularly well across consumers and baseline scenarios. In comparison, the secondary consumer in scenario 3 of the bulk stable isotope model had an estimation error of 0.8 trophic levels (Figure \@ref(fig:BSIATP)C) which was twice as high as the largest trophic position estimation error for the CSIA model.

All amino acids used in the CSIA model performed better in terms of the magnitude of trophic position estimation error (Figure \@ref(fig:CSIATP)) and duration of deviation compared to BSIA analysis (Figure \@ref(fig:BSIATP)). Of the tested trophic amino acids, proline had the greatest magnitude of estimation error for both shrimp, the primary consumer (maximum estimation error = 0.4, Figure \@ref(fig:CSIATP)G), and tuna, the secondary consumer (maximum estimation error = 0.4; Figure \@ref(fig:CSIATP)F). In comparison, the canonically used trophic amino acid for CSIA trophic position calculations, glutamic acid, only had a maximum estimation error of 0.04 for shrimp and 0.2 for tuna for the same baseline, and lower magnitude of estimation error for all other baselines (Figure \@ref(fig:CSIATP)E). Both glutamic acid and alanine performed similarly in terms of estimation error for both consumers in the CSIA model and as a result were the best suited trophic amino acids for minimizing trophic position estimation error when paired with phenylalanine and lysine as source amino acids.

###Sensitivity Analysis

Sensitivity of trophic position estimation error to consumer half-life was similar across consumer trophic level for BSIA of nitrogen. For all consumers, the mean and standard deviation of trophic position estimation error was similar at half-life values greater than 100 days. For higher trophic level consumers, trophic position was slightly more sensitive to half-life values less than 100 days (Figure \@ref(fig:SensBSIA)C & D). Initially the deviation and variability increased at a higher rate than consumers lower in the food web, but eventually leveled off at a half-life of approximately 100 days, at which point the mean and standard deviation of the tertiary consumer was similar to that of lower trophic level consumers (Figure \@ref(fig:SensBSIA)D).

Trophic position deviation for BSIA was more sensitive to consumer half-life than CSIA. Sensitivity of trophic position deviation was minimized when the half-life of the trophic amino acid was equal to the half-life of the source amino acid, which was 130 days (Figure \@ref(fig:SensCSIAlys)) and 33 days (Figure \@ref(fig:SensCSIAphe)) for these model simulations. Mean trophic position deviation and its distribution decreased in magnitude initially before intersecting with the half-life of source amino acid. At half-life value greater than the source amino acid variability increased slightly but did not deviate substantially from zero for any of the three trophic level consumers (Figure \@ref(fig:SensCSIAlys)D). The greatest sensitivity of tropic position estimation error occurs when the half-life values of trophic amino acids are less than source amino acid half-life values. 

###Comparison to recommended methodologies

Alternative recommended methods to calculate trophic position from stable isotope data produced only moderate improvements for trophic position accuracy. The 90-day lag between the stable isotope baseline and tertiary consumers improved accuracy for trophic position, although estimates ranged ± 0.6 trophic levels for the eutrophic, arctic lake (Table \@ref(tab:rangetp)A, Scenario 4). The range in trophic position estimates for secondary consumers were similar for the 90-day lag compared to a simultaneous sampling approach. In contrast the range in primary consumer trophic position was much greater with a 90-day lagged sampling strategy compared to a simultaneous strategy, with overestimates as high as two trophic levels in some systems (Table \@ref(tab:rangetp)A). Using primary consumers as a stable isotope baseline improved trophic position estimation error for all consumers in all systems (Figure \@ref(fig:BSIATP2)) compared to the particulate organic isotope baseline model (Figure \@ref(fig:BSIATP)) although estimation error still exceeded ± 0.5 for substantial amounts of time for most scenarios. Applying the 90-day lagged equation with the primary consumer as the stable isotope baseline further improved trophic position accuracy for the tertiary consumer (Table \@ref(tab:rangetp)B) although accuracy for the secondary consumer was similar to the simultaneously measured model with the primary consumer as the stable isotope baseline (Figure \@ref(fig:BSIATP2)).

##Discussion

Temporal heterogeneity in the bulk nitrogen stable isotope baseline results in erroneous trophic position estimates when heterogeneity and tissue turnover are not accounted for. Based on our results, this error can even occur when the isotope baseline can be sampled continuously and simultaneously with the consumer and is considered known. Both the shape of the heterogeneity of the stable isotope baseline and the trophic level of the consumers are important factors in the magnitude and duration of trophic position estimation error for a given aquatic food web. Notably, for BSIA, trophic position estimation error propagates up the food web with greater error in estimation for consumers feeding at higher trophic levels. Using CSIA to sample consumers provides the best approach for minimizing trophic position estimation error due to temporal heterogeneity and tissue turnover rates however the relative rate of turnover between the source amino acid and the trophic amino acid used to calculate trophic position should be considered. When CSIA is not feasible, using primary consumers as the stable isotope baseline or sampling the consumer 90 days after sampling the baseline can provide moderate improvement to trophic position estimation error of high trophic level consumers, but does not eliminate it entirely and the improvement comes with decreased accuracy for lower trophic level consumers.

The duration of trophic position deviation is substantial for all of the modelled consumers using BSIA. However, no consumer had a trophic position deviation of greater than ± 0.5 for the entire modelled time period. In fact, for at least 30% of the model time for all consumers in all food webs trophic deviation was less than ± 0.5 trophic positions. Based on this result, there are substantial time periods within each food web when trophic deviation can be minimized simply based on sampling time. For most modelled food webs, days 0 - 100 (January 1 – April 10) and 200 - 250 (July 19 – September 7) produced accurate trophic position estimations for secondary and tertiary consumers (Figure \@ref(fig:BSIATP) A-F), although the moderately eutrophic, urban lake (scenario 5; Figure \@ref(fig:BSIATP)E) was an exception. Sampling consumers during early winter or late summer when estimation error is lowest could improve accuracy of trophic position estimation. Consideration for the shape of the stable isotope baseline is necessary for this approach. Specifically, researchers should identify how nitrogen resources are expected to change through time in their system of interest and anticipate how this may impact the stable isotope values of consumers. However, this approach may not be appropriate for all research questions, as trophic position measurements during the spring or autumn may be of particular interest. The applicability of this strategy across systems also warrants additional research as the food web baselines modelled in this study represent temperate systems.

Notably, estimation errors associated with turnover and heterogeneity propagate up food webs and are more pronounced and occur for a longer duration (Table \@ref(tab:durdev)) in higher trophic level consumers when using BSIA to estimate trophic position (Figure \@ref(fig:BSIATP)). The sensitivity of trophic position to stable isotope half-life also propagates up the food web (Figure \@ref(fig:SensBSIA)), such that higher trophic level consumers experience more trophic position deviation at higher half-life values than their lower trophic level counterparts. This is a result of tertiary consumers being even further from isotopic steady state with the sampled isotope baseline, in this case POM, compared to lower trophic level consumers. @Cabana1996 observed that $\delta^{15}N$ values of phytoplankton are ten times more variable than primary consumers. Due to the short life span of primary producers, they only represent a short-term measure of the isotope baseline [@Vizzini2003]. By sampling primary consumers as the stable isotope baseline isotope heterogeneity is time averaged as nitrogen is assimilated into primary consumer tissues. This results in a dampening of baseline stable isotope heterogeneity. Our results agree with this, while trophic position estimation error is higher in secondary and tertiary consumers, variability in the magnitude of trophic position estimation error is dampened relative to the stable isotope values observed lower in the food web (Figure \@ref(fig:BSIA15N)). This indicates that when consumer tissues are sampled is less impactful for the overall magnitude of trophic position deviation than when and how the baseline is sampled. Modifications to the timing of stable isotope baseline sampling is the most fruitful approach for minimizing trophic position estimation error (compared to adjusting consumer sampling).

Previous research has provided guidance to improve accuracy of trophic position estimation from bulk stable isotope data. Our analysis utilized stable isotope data of particulate organic matter because it is prevalent and longer-term data was available for a greater variety of systems. It also allowed us to manipulate the tissue turnover time of all of the consumers in the modelled food web. However, @Kjeldgaard2021 found that there were 10 methods to quantify stable isotope baseline in the literature, of which particulate organic matter was only one. Guidance on stable isotope sampling strategy has included increasing sample size [@Kjeldgaard2021], using primary consumers as the stable isotope baseline rather than particulate organic matter or primary producers [@Post2002], and sampling consumers of interest and the stable isotope baseline 90 days apart [@Possamai2021]. With our model, we explored these options and found that applying a 90-day lag or using the primary consumer as the stable isotope baseline offered only moderate improvement of the accuracy in trophic position estimation (Table \@ref(tab:rangetp); Figure \@ref(fig:BSIATP2)). Accuracy in trophic position estimation improved the most for tertiary consumers for both methods but was worse for primary consumers when applying the 90-day delay. This result indicates the  steady-state assumption is more accurate for higher trophic level consumers when applying a 90-day sampling delay and using primary consumers as the stable isotope baseline. When applying both of the sampling strategies at once, the trophic position deviation of tertiary consumers was only ± 0.3 trophic levels for most systems (Figure \@ref(fig:BSIA15N)B). Using a 90-day lag between the consumer of interest and isotope baseline, using primary consumers as the stable isotope baseline, or both, could be a useful strategy for minimizing error in trophic position estimation for tertiary consumers or higher but will decrease the accuracy of lower trophic level consumers.

To further improve this strategy, a different time lag could be applied for lower trophic level consumers that is less than 90 days and aligns more closely with their isotopic turnover rates. This strategy would be particularly useful for studies where large sample size or CSIA is impractical or cost prohibitive. It is important to note that in order to utilize primary consumers as the isotopic baseline as opposed to primary producers, the trophic level of the primary consumers must be known with certainty, or it could introduce another component of error into the trophic position estimates. This can be particularly challenging when employing a method such as plankton tows, which frequently include mixture of planktonic taxa (i.e., microbial heterotrophs, detritus, microzooplankton) in different life stages, and as a result can impact trophic position estimates of consumers of any trophic level [@McCarthy2007; @Hannides2013]. In addition, omnivory is common in zooplankton [@Atkinson1996; @Castellani2008] making it challenging to know their trophic level absolutely if they are used as the stable isotope baseline for trophic position calculations of other consumers. Using sessile filter feeders (i.e. mussels, barnacles) as the stable isotope baseline when possible would be the best practice, as suggested by other research [@VanderZanden1999; @VanderZanden2001; @Post2002; @Mancinelli2013].

CSIA substantially reduces the trophic position estimation error and duration of trophic position deviation relative to BSIA. Some trophic amino acids performed better than others, specifically, proline had the highest trophic position estimation error compared to glutamic acid and alanine which had the lowest. Based on our sensitivity analysis, trophic position estimation error is less sensitive to longer trophic amino acid half-life values (Figures \@ref(fig:SensCSIAlys) & \@ref(fig:SensCSIAphe)). Calculating trophic position using a source amino acid with a longer half-life (Figure \@ref(fig:SensCSIAlys)) also reduced sensitivity of estimation error to trophic amino acid half-life values compared to a shorter half-life of the source amino acid (Figure \@ref(fig:SensCSIAphe)). 

CSIA allows researchers to measure the stable isotope value of multiple amino acids from a single sample. Phenylalanine and glutamic acid have become the canonical source and trophic amino acids applied in CSIA because phenylamine demonstrates the lowest trophic enrichment while glutamic acid has the highest [@McClelland2002; @Chikaraishi2007; @Chikaraishi2009; @Hannides2009]. Recent studies have shown using multiple amino acids can improve trophic position estimates (@Nielsen2015, Feddern et al. in prep) but there is little guidance in the literature quantifying which amino acids are best and why. Best practices in CSIA should focus on utilizing source - trophic amino acid pairs where the amino acids have similar half-life values. When researchers are choosing between pairs, it is best to select source amino acids with longer half-life values, and generally, selecting a trophic amino acid that has a longer half-life than the selected source amino acid will produce the best results. Based on our results, glutamic acid and alanine are the best trophic amino acids to use and are suitable to be paired with either lysine or phenylalanine as source amino acids.

While our results show CSIA reduces error in trophic position estimation relative to BSIA, CSIA is can be cost prohibitive and half-life values of individual amino acids are not well studied across taxa and tissues. Studies examining nitrogen stable isotope half-life values of individual amino acids are limited, and the field of CSIA would greatly benefit from more half-life values reported in the literature. Our results show lysine, phenylalanine, glutamic acid, and alanine perform well based on half-life values of tuna white muscle and shrimp tail muscle. Similarly, Feddern et al. in prep found that using both alanine and glutamic acid paired with phenylalanine from harbor seal bone offered the most improvement to model certainty compared to using either amino acid alone or other trophic amino acids. This indicates that the best CSIA amino acid pairs may be fairly consistent across taxa and tissues, but additional studies of tissue turnover rates would verify this result, particularly in animals that utilize urea-excretion [@Germain2013] and in tissues commonly used for historical CSIA studies such as bone or coral [@Misarti2009; @Feddern2021; @Sherwood2005; @Sherwood2008; @Sherwood2011].

Here we focused on seasonal, temporal heterogeneity, but other types of stable isotope heterogeneity exist within aquatic systems and impact trophic position calculations. Spatial heterogeneity is common in stable isotope baselines [@Lorrain2015] and changes with latitude [@Lorrain2015; @Vokshoori2014] and depth [@Wu1999; @Sigman2009; @Peters2012] within ocean basins. As a result, mobile consumers are frequently integrating distinct stable isotope baselines, that may not reflect the baseline where they were caught for sampling. Stable isotope values of primary producers also vary based on taxa in both freshwater [@Vuorio2006] and marine habitats [@Ramirez2021; @Lorrain2015; @Sigman2009]. On smaller scales in freshwater lakes, habitat availability and population abundance can also influence isotopically distinct basal resource use of aquatic consumers [@Stiling2021]. Changes in diet that impact resource use can result in changes in the stable isotope baseline that is assimilated by a consumer of interest. Similar considerations for tissue turnover times and its implications for trophic position estimates should be made for mobile consumers in spatially heterogenous systems as well as the temporally heterogenous systems examined in this study.

This study examined temporal isotope heterogeneity, tissue turnover times and their impact of estimating trophic position from stable isotope data. While we focused on the assumption of isotopic steady state between consumers and their prey, this is not the only assumption for calculating stable isotope values that impacts the interpretation of stable isotope data. Trophic enrichment factors are often assumed to be equal for consumers but is known to be vary based on taxa, tissues, excretion pathway, and diet [@Deniro1981; @Vanderklift2003; @Caut2008; @Blanke2017].  We did not model the effect of variable trophic enrichment factors, though this value is known to influence trophic position calculations and resource estimates [@Bond2011]. Variability in trophic enrichment factors and baseline heterogeneity can act both synergistically and antagonistically on trophic position estimates. The scope of this study was to understand error associated with heterogeneity and tissue turnover specifically. While our model is limited in the number of parameters that vary, it fulfills this crucial knowledge gap. 

Stable isotope heterogeneity and stable isotope turnover of consumer tissues impact the accuracy of trophic position estimates. Based on our results CSIA reduces trophic position estimation error more than any of the tested trophic position equations for BSIA. Using primary consumers as the stable isotope baseline or applying a 90-day lag between baseline and consumer sampling can improve accuracy of trophic position estimation for tertiary consumers or higher, although it does not eliminate it. Thus, stable isotope heterogeneity and its potential effects on trophic position estimates should be considered when interpreting stable isotope values. Our results agree with the recommendations made by @Possamai2021), when possible, trophic position studies should use CSIA for at least a subset of samples. Researchers applying stable isotope data to calculate consumer trophic position should carefully consider heterogeneity in baseline isotope values expected in their system, isotope turnover rates for their consumers of interest, and isotope turnover rates of the resources of their consumers of interest. We caution researchers against a one size fits all approach for improving trophic position estimation, and it may be beneficial to apply different approaches for different consumers in the same system.

\clearpage

##Tables

**Table** \@ref(tab:base): Summary of observational nitrogen stable isotope baseline data used for modelling case studies.  

```{r base,echo=FALSE}
supp1 <- read.csv("data/Ch5/Table1.csv")
colnames(supp1) <- c("Study",
                     "Taxa",
                     "Season of Data",
                     "Range of \n Isotopic Change",
                     "Ecosystem",
                     "Location")
supp1 <- kableExtra:::kable_styling(knitr::kable(supp1, "latex", escape = FALSE,
caption = "Baseline environments",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8, full_width=F)
 
kableExtra::column_spec(supp1, column=1:6, width = "8em")

```
\clearpage

**Table** \@ref(tab:durdev):  Duration of trophic position deviation (number of days the magnitude of trophic position estimation error is greater than ± 0.5 trophic levels from ($\hat{tp}$) for each of the of the six, bulk stable isotope food web scenarios (Figure 5.1).
```{r durdev,echo=FALSE}
supp1 <- read.csv("data/Ch5/Table2.csv")
colnames(supp1) <- c("Trophic Position",
                     "Scenario 1",
                     "Scenario 2",
                     "Scenario 3",
                     "Scenario 4",
                     "Scenario 5",
                     "Scenario 6")
supp1 <- kableExtra:::kable_styling(knitr::kable(supp1, "latex", escape = FALSE,
caption = "Trophic Position Duration of Deviation",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8)
kableExtra::column_spec(supp1, column=1:5, width = "6em")
```
\clearpage

**Table** \@ref(tab:dur): Duration of trophic position deviation (number of days the magnitude of trophic position estimation error is greater than ± 0.5 trophic levels from ($\hat{tp}$) for each of the of the two CSIA food web scenarios (Figure \@ref(fig:CSIA)).

```{r dur,echo=FALSE}
supp1 <- read.csv("data/Ch5/Table3.csv")
colnames(supp1) <- c("Amino Acid",
                     "Baseline 1 (365)",
                     "Baseline 2 (365)",
                     "Baseline 3 (233)",
                     "Baseline 4 (351)")
supp1 <- kableExtra:::kable_styling(knitr::kable(supp1, "latex", escape = FALSE,
caption = "Duration of Trophic Position Deviation (days)",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8)
kableExtra::column_spec(supp1, column=1, width = "8em")
```
\clearpage

**Table** \@ref(tab:rangetp): Range of trophic position estimates applying a 90-day lag between baseline sampling and consumer sampling for calculations A) trophic position estimates using POM as the stable isotope baseline and B) trophic position estimates using the primary consumer as the stable isotope baseline

```{r rangetp,echo=FALSE}
supp1 <- read.csv("data/Ch5/Table4.csv")
colnames(supp1) <- c("Scenario",
                     "Baseline",
                     "Primary Consumer (tp = 2)",
                     "Secondary Consumer (tp = 3)",
                     "Tertiary Consumer (tp = 4)")
supp1 <- kableExtra:::kable_styling(knitr::kable(supp1, "latex", escape = FALSE,
caption = "Range of Trophic position estimates with different baselines",
                  longtable = TRUE, booktabs = TRUE
                  ),font_size = 8)
kableExtra::column_spec(supp1, column=1:5, width = "6em")
```
\clearpage

##Figures
\begin{landscape}
Figure 5.1: Six theoretical food web scenarios for BSIA model constructed from consumer experimental stable isotope half-life ($t_{0.5}$) studies and observed nitrogen stable isotopes values of particulate organic matter (baselines 1-4; Table 5.2) from four distinct aquatic systems. Three comparisons were made between two food web scenarios: 1) different $t_{0.5}$ of the secondary consumer 2) different $t_{0.5}$ of the secondary and tertiary consumer 3) different baseline from particulate organic matter. 
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=1.05\textwidth]{figure/Ch5/Figure1.pdf}
  \caption{BSIA theoretical food web scenarios}
  (\#fig:BSIA)
\end{figure}
\end{landscape}
\clearpage

**Figure** \@ref(fig:CSIA): A theoretical food web for CSIA model constructed from consumer experimental stable isotope half-life ($t_{0.5}$) studies and applied to observed nitrogen stable isotopes values of particulate organic matter from all four distinct aquatic systems (baselines 1- 4, Table 5.2).
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.75\textwidth]{figure/Ch5/Figure2.pdf}
  \caption{CSIA theoretical food web scenarios}
  (\#fig:CSIA)
\end{figure}
\clearpage

**Figure** \@ref(fig:EffectSize): Depiction of the two effect sizes used to assess accuracy of trophic position estimation in this study 1) the magnitude of trophic position estimation error which describes the degree to which trophic position is over or under estimated and 2) the duration of deviation which describes the length of time over the course of the model simulation that trophic position was erroneously estimated by ± 0.5 trophic levels. 
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.85\textwidth]{figure/Ch5/Figure3.pdf}
  \caption{Trophic position estimation error effect sizes}
  (\#fig:EffectSize)
\end{figure}
\clearpage

**Figure** \@ref(fig:BSIA15N): Simulated primary, secondary, and tertiary consumer bulk nitrogen stable isotope values ($\delta^{15}N$) for six (A-F) food web scenarios (Figure 5.1) using $\delta^{15}N$ of particulate organic matter from observational studies as the stable isotope baseline (Table 5.1).  
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.65\textwidth]{figure/Ch5/Figure4.pdf}
  \caption{BSIA simulted $\delta^{15}N$}
  (\#fig:BSIA15N)
\end{figure}
\clearpage

**Figure** \@ref(fig:BSIATP): Trophic position estimation error based on simulated primary, secondary, and tertiary consumer bulk nitrogen stable isotope values ($\delta^{15}N$) for six (A-F) food web scenarios (Figure 5.1) and three main comparisons using POM stable isotope values to calculate trophic position. 0 denotes the true trophic level of a given consumer and grey box denotes trophic position deviation of ± 0.5 trophic levels and used to calculate duration of deviation.
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.6\textwidth]{figure/Ch5/Figure5.pdf}
  \caption{BSIA simulated trophic position}
  (\#fig:BSIATP)
\end{figure}
\clearpage

**Figure** \@ref(fig:CSIA15N): Simulated amino acid (color) nitrogen CSIA values ($\delta^{15}N$) for primary (A, C, E, G) and secondary (B, D, F, H) consumers using four stable isotope baselines derived from observational particulate organic matter studies (Table 5.1, Figure \@ref(fig:CSIA)).
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.6\textwidth]{figure/Ch5/Figure6.pdf}
  \caption{CSIA simulated $\delta^{15}N$}
  (\#fig:CSIA15N)
\end{figure}
\clearpage

**Figure** \@ref(fig:CSIATP): Trophic position estimation error from simulated amino acid (color) nitrogen CSIA values ($\delta^{15}N$) for primary (A, C, E, G) and secondary (B, D, F, H) consumers using four stable isotope baselines derived from observational particulate organic matter studies (Table 1, Figure \@ref(fig:CSIA)). 0 denotes the true trophic level of a given consumer and grey box denotes trophic position deviation of ± 0.5 trophic levels and used to calculate duration of deviation.
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.55\textwidth]{figure/Ch5/Figure7.pdf}
  \caption{CSIA simulated trophic position}
  (\#fig:CSIATP)
\end{figure}
\clearpage

**Figure** \@ref(fig:SensBSIA): Sensitivity analysis of the magnitude of trophic position estimation error from bulk stable isotope values in response to primary, secondary, and tertiary (color) consumer stable isotope half-life values for a single simulated stable isotope baseline (a). Solid line denotes mean magnitude of estimation error across the 125-day baseline simulation (A) for a given half-life and shaded region denotes 1 standard deviation from the mean. 0, in B-D, represents the true trophic level of a consumer.
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.55\textwidth]{figure/Ch5/Figure8.pdf}
  \caption{BSIA tissue turnover sensitivity analysis}
  (\#fig:SensBSIA)
\end{figure}
\clearpage

**Figure** \@ref(fig:SensCSIAlys): Sensitivity analysis of the magnitude of trophic position estimation error from CSIA for primary, secondary, and tertiary (color) consumer stable isotope half-life values for a single simulated stable isotope baseline (a). The source amino acid line denotes the single half-life value (130 days, lysine tuna) of the source amino acid modelled in this analysis. Solid line denotes mean magnitude of estimation error across the 125-day baseline simulation (A) for a given half-life and shaded region denotes 1 standard deviation from the mean. 0, in B-D, represents the true trophic level of a consumer.
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.75\textwidth]{figure/Ch5/Figure9.pdf}
  \caption{CSIA tissue turnover sensitivity analysis: lysine}
  (\#fig:SensCSIAlys)
\end{figure}

\clearpage

**Figure** \@ref(fig:SensCSIAphe): Sensitivity analysis of the magnitude of trophic position estimation error from CSIA for primary, secondary, and tertiary (color) consumer stable isotope half-life values for a single simulated stable isotope baseline (a). The source amino acid line denotes the single half-life value (33 days, phenylamine in shrimp) of the source amino acid modelled in this analysis. Solid line denotes mean magnitude of estimation error across the 125-day baseline simulation (A) for a given half-life and shaded region denotes 1 standard deviation from the mean. 0, in B-D, represents the true trophic level of a consumer.
\newline
\begin{figure}[h]
\centering
  \includegraphics[width=0.75\textwidth]{figure/Ch5/Figure10.pdf}
  \caption{CSIA tissue turnover sensitivity analysis: phenylalanine}
  (\#fig:SensCSIAphe)
\end{figure}
\clearpage

**Figure** \@ref(fig:TissueTurn): Nitrogen stable isotope half-life values (days) experimentally derived for a) bulk stable isotope analysis [@VanderZanden2015] for different species (label) and tissues (color) and b) CSIA half-life values of individual amino acids (Downs et al. and Popp et al.) for different species (color) and amino acids (label). 
\newline
\begin{figure}[h]
\centering
  \includegraphics[width=0.85\textwidth]{figure/Ch5/Figure11.pdf}
  \caption{Observed tissue half-life values}
  (\#fig:TissueTurn)
\end{figure}
\clearpage

**Figure** \@ref(fig:BSIATP2): Trophic position estimation error based on simulated primary, secondary, and tertiary consumer bulk nitrogen stable isotope values ($\delta^{15}N$) for six (A-F) food web scenarios (Figure 5.1) using the primary consumer stable isotope values to calculate trophic position. 0 denotes the true trophic level of a given consumer and grey box denotes trophic position deviation of ± 0.5 trophic levels used to calculate duration of deviation.
\newline 
\begin{figure}[h]
\centering
  \includegraphics[width=0.6\textwidth]{figure/Ch5/Figure12.pdf}
  \caption{BSIA simulated trophic position: primary consumer baseline}
  (\#fig:BSIATP2)
\end{figure}
\clearpage
